Harnessing the cloud to understand light: A whitepaper detailing the \MaxwellFDS core solver

### 2013-05-24 thesis/bibliography.tex
Thesis revision (from Jelena's comments) almost done.
Just need conclusion.

@@

\bibitem{Lu12} J. Lu, J. Vuckovic, ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes,'' 
    Opt. Express \textbf{20}, 7221-7236 (2012)

\bibitem{Lu13} J. Lu, J. Vuckovic, ``Nanophotonic computational design,'' \emph{under review}.

@@

>>>>

\bibitem{pic1} 
    Qianfan Xu, Bradley Schmidt, Sameer Pradhan, and Michal Lipson,
    ``Micrometre-scale silicon electro-optic modulator,''
    Nature \textbf{435}, 325-327 (2005).

\bibitem{pic2}
    Jie Sun, Erman Timurdogan, Ami Yaacobi, Ehsan Shah Hosseini, and Michael R. Watts,
    ``Large-scale nanophotonic phased array,''
    Nature \textbf{493}, 195-199 (2013).

\bibitem{lipson} 
    A. Gondarenko and M. Lipson, 
    ``Low modal volume dipole-like dielectric slab resonator,'' 
    Opt. Express \textbf{16}, 17689-17694 (2008).

\bibitem{yablo} C.-Y. Kao, S. Osher, and E. Yablonovitch, 
    ``Maximizing band gaps in two-dimensional photonic crystals by using level set methods,'' Appl. Phys. B \textbf{81}, 235-244 (2005).

\bibitem{levi}
    P. Seliger, M. Mahvash, C. Wang, and A. Levi,
    ''Optimization of aperiodic dielectric structures,''
    J. Appl. Phys. \textbf{100}, 34310-6 (2006).

\bibitem{johnson} A. Oskooi, A. Mutapcic, S. Noda, J. D. Joannopoulos, S. P. Boyd, and S. G. Johnson, 
    ``Robust optimization of adiabatic tapers for coupling to slow-light photonic-crystal waveguides,'' Opt. Express \textbf{20}, 21558-21575 (2012).

\bibitem{sigmund1} Y. Elesin, B. S. Lazarov, J. S. Jensen, O. Sigmund, 
    ``Design of robust and efficient photonic switches using topology optimization,''
    Photonics and Nanostructures - Fundamentals and Applications \textbf{10}, 
    153–165 (2012).

\bibitem{jameson} L. Martinelli and A. Jameson, 
    ``Computational aerodynamics: solvers and shape optimization,''
    J. Heat Transfer \textbf{135}, 011002 (2013).

\bibitem{bendsoe} M. P. Bendsoe and O. Sigmund,
    ``Material interpolation schemes in topology optimization,''
    Archive of Applied Mechanics \textbf{69}, 635-654 (1999).

@@

\bibitem{Lu12} J. Lu, J. Vuckovic, ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes,'' 
    Opt. Express \textbf{20}, 7221-7236 (2012).

\bibitem{Lu13} J. Lu, J. Vuckovic, ``Nanophotonic computational design,'' 
    Opt. Express \textbf{21}, 13351-13367 (2013).

@@

\bibitem{robust1} F. Wang, J. S. Jensen, and O. Sigmund, ``Robust topology optimization of photonic crystal waveguides with tailored dispersion properties,'' J. Opt. Soc. Am. B \textbf{28}, 387-397 (2011).

### 2013-05-24 thesis/chapter1.tex
Thesis revision (from Jelena's comments) almost done.
Just need conclusion.

\section{Problem formulation}
This work was originally sparked by the realization that 
    the electromagnetic wave equation was not only linear in the field variable,
    but also in the structure variable.
Specifically, that
    \BE \nabla \times \mu_0^{-1} \nabla \times E - \omega^2 \epsilon E = 
        -i\omega J \label{p1:wave} \EE
    is \emph{bi-linear} in both $E$ and $\epsilon$.
Practically, this meant that solving for $\epsilon$ should be rather straightforward,
    since we already now how to solve linear systems in general.

@@

\chapter{Regularized direct design}\label{regularized}
\section{Problem formulation}
We were previously quite excited to learn that solving for 
    a dielectric structure ($\epsilon$ or $z$) given 
    its electromagnetic field ($E$ or $x$) is actually quite simple,
    but generally results in unrealizable structures.
Along these lines, we also knew that the simplest way 
    to produce a well-behaved dielectric structure is 
    to add a regularization term to our direct formulation (\eq{p1:axb}),
    \BE \minimize_z \|A(z)x - b\|^2 + \eta \|z - z_0\|^2. \label{reg:problem}\EE
Here, the regularization term $\eta \|z - z_0\|^2$
    serves to keep $z$ from straying too far from $z_0$, which we choose.
The coefficient $\eta$, then, serves as a knob to control the
    strength of the regularization term.

@@

Although our results seemed promising in producing well-behaved
    dielectric structures;
    critically, fig.~\ref{regls pic} shows that there seems
    to be an unavoidable trade-off between 
    constraining the structure ($z$) 
    and accurately reproducing the field ($x$).

>>>>

\section{Problem formulation}
This work was originally sparked by the realization that 
    the electromagnetic wave equation was not only linear in the field variable,
    but also in the structure variable.
Specifically, that both $E$ and $\epsilon$ are \emph{bi-linear} in
    the electromagnetic wave equation which is
    \BE \nabla \times \mu_0^{-1} \nabla \times E - \omega^2 \epsilon E = 
        -i\omega J, \label{p1:wave} \EE
    where
    \BI $E$ is the electric-field,
    \I  $J$ is the excitation current density,
    \I  $\omega$ is the angular frequency,
    \I  $\mu_0$ is the permeability of vacuum, and
    \I  $\epsilon$ is the space-dependent permittivity. \EI
Practically, this meant that solving for $\epsilon$ should be rather straightforward,
    since we already now how to solve linear systems in general.

@@

\chapter{Regularized direct design}\label{regularized}
\section{Problem formulation}
We were previously quite excited to learn that solving for 
    a dielectric structure ($\epsilon$, i.e.~$z$) given 
    its electromagnetic field ($E$, i.e.~$x$) is actually quite simple,
    but generally results in unrealizable structures.
Along these lines, we also knew that the simplest way 
    to produce a well-behaved dielectric structure is 
    to add a regularization term to our direct formulation (\eq{p1:axb}),
    \BE \minimize_z \|A(z)x - b\|^2 + \eta \|z - z_0\|^2. \label{reg:problem}\EE
Here, the regularization term $\eta \|z - z_0\|^2$
    serves to keep $z$ from straying too far from $z_0$, which we choose.
The coefficient $\eta$, then, serves as a knob to control the
    strength of the regularization term.

@@

Although our results seemed promising in producing well-behaved
    dielectric structures,
    fig.~\ref{regls pic} shows that there seems
    to be an unavoidable trade-off between 
    constraining the structure ($z$) 
    and accurately reproducing the field ($x$).

### 2013-05-24 thesis/chapter2.tex
Thesis revision (from Jelena's comments) almost done.
Just need conclusion.

Further thought led us to conclude that
    the fundamental problem in the examples 
    in chapters \ref{direct} and \ref{regularized}
    is actually not in the methods themselves, 
    but in the improper selection of a target field. 
In fact, we realizaed thtat it is very difficult to select a suitable target field 
    because not every field even has a manufacturable dielectric structure 
    that is able to reproduce it. 
Furthermore, it is nearly impossible to select a multi-dimensional field 
    which corresponds to a well-behaved, isotropic and 
    discretely-valued $\epsilon$, 
    as would be needed for practical structures.

@@

\section{Results}
\subsection{1D}
Fig.~\ref{comp pic} shows that the iterative algorithm, after $400$ iterations and with the correct choice of regularization parameters $\eta_0$ and $\eta_1$, 
    indeed results in a well-behaved structure that is able to closely reproduce the modified target field\cite{Lu10}!

@@

\subsection{2D}
The iterative design method can also be extended to two dimensions\cite{Lu10}.
We demonstrate this using an S-shaped target field 
    which is non-trivial to reproduce. 
The design result is shown in fig.~\ref{S pic}. 
    The resulting dielectric structure is continuous, unbounded and 
    contains very few singularities (white dots), 
    but the final target and actual fields match up well. 
Also, the computational cost remains quite reasonable; the $50$ iterations needed required only $5$ minutes of computation time. 
\begin{figure}\centering
\includegraphics[width=\textwidth]{p1/S}
\caption{Iterative design of an ``S'' resonator. 
The design was initialized by specifying an initial dielectric structure ($\epsilon=1$ everywhere) and a resonant field in the shape of an ``S''. 
The final dielectric structure was produced after $50$ iterations which took $90$ seconds to complete in total. The grid size was $80\times 120$. The final dielectric structure is quite unintuive, and yet reproduces the target field surprisingly well. This example demonstrates the versatility of the iterative design method in producing designs, from very simple specifications, which otherwise could be attained only with considerable difficulty.}
\label{S pic}
\end{figure}

@@

% 
% \documentclass[10pt,letterpaper]{article}
% \usepackage{opex3}
% \usepackage{amsmath}
% \usepackage{graphicx}
% 
% \begin{document}
% \title{Inverse design of a three-dimensional nanophotonic resonator}
% \author{Jesse Lu, Stephen Boyd, and Jelena Vu\v{c}kovi\'{c}}
% \address{Stanford University, \\ Stanford, C.A., 94305}
% \email{jesselu@stanford.edu}
% 
% 
% \begin{abstract} 
% The inverse design of a three-dimensional nanophotonic resonator is presented. The design methodology is computationally fast (10 minutes on a standard desktop workstation) and utilizes a 2.5-dimensional approximation of the full three-dimensional structure. As an example, we employ the proposed method to design a resonator which exhibits a mode volume of $0.32(\lambda/n)^3$ and a quality factor of $7063$. % 8808 35648
% \end{abstract}
% 
% \ocis{(230.5750) Resonators.} % REPLACE WITH CORRECT OCIS CODES FOR YOUR ARTICLE
% 
% %%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{thebibliography}{99}
% \bibitem{miller} D. A. B. Miller, ``Rationale and challenges for optical interconnects to electronic chips,'' Proc. of the IEEE \textbf{88}, 728-749 (2000).
% \bibitem{prevwork} J. Lu and J. Vuckovic, ``Inverse design of nanophotonic structures using complementary convex optimization,'' Opt. Express \textbf{18}, 3793-3804 (2010).
% \bibitem{inan} U. Inan, A. Inan, \emph{Electromagnetic Waves} (Prentice Hall, 2000), page 296.
% \bibitem{yee} K. Yee, ``Numerical solution of initial boundary value problems involving maxwell's equations in isotropic media,'' IEEE Trans. Antennas Propag. Mag. \textbf{14}, 302-307 (1966).
% \bibitem{boydbook} S. Boyd and L. Vandenberghe, \emph{Convex Optimization} (Cambridge University Press, 2004).
% \bibitem{altdir} S. Boyd, N. Parikh, E. Chu, B. Peleato, and J. Eckstein are preparing a manuscript to be called, ``Distributed Optimization and Statistical Learning via the Alternating Direction Method of Multipliers,'' \verb+http://www.stanford.edu/~boyd/papers/distr_opt_stat_learning_admm.html+. 
% \bibitem{cholmod} Y. Chen, T. A. Davis, W. W. Hager, and S. Rajamanickam, ``Algorithm 887: CHOLMOD, supernodal sparse Cholesky factorization and update/downdate,'' ACM Trans. Math. Software \textbf{35}, No. 3, 2009.
% \bibitem{cvx} M. Grant and S. Boyd, \emph{CVX: Matlab software for disciplined convex programming}, version 1.21. \verb+http://cvxr.com/cvx+, January 2011.
% \bibitem{digitize} D. Englund, I. Fushman, and J. Vuckovic, ``General recipe for designing photonic crystal cavities,'' Opt. Express \textbf{13}, 5961-5975 (2005).
% \end{thebibliography}
%  
% %%%%%%%%%%%%%%%%%%%%%%%%%%  body  %%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Begin of 2.5 D (Motivation)}
% To date, the design of nanophotonic devices has generally involved a lengthy (days, weeks) process in which one perturbs, by trial-and-error, canonical structures such as photonic crystals or waveguide-coupled rings to achieve the desired performance for the device. 
% 
% Instead, an inverse design method in which the user only specifies the desired electromagnetic field, or characteristics thereof, and then leaves the computer to find a dielectric structure satisfying these requirements, would be a much more intuitive and computationally efficient design strategy. Furthermore, such a method may even be the only feasible option available for the design of more complex multi-wavelength and multi-functional nanophotonic devices needed for on-chip integration\cite{miller}.
% 
\subsection{2.5D}
% \section{2.5-dimensional approximation}
Unfortunately, to directly extend our previous method to full three-dimensional space requires solving for a very large ($\sim 10^7 \times 10^7$) and ill-conditioned matrix, namely that given by the time-harmonic Maxwell's equation in three dimensions, which for the electric ($E$) field is
\begin{align}
\nabla\times\nabla\times E - \mu\omega^2\epsilon E &= 0,\quad\text{where}\label{wave} \\
E &= E(x,y,z).
\end{align}

@@

In this approximation, converting from a 2.5-dimensional structure to a full three-dimensional structure only requires the correct choice of slab thickness, $t$, since the in-plane values of $\epsilon$ remain the same. To estimate $t$ for the full three-dimensional device, we treat the resonant mode as a guided mode in a slab waveguide surrounded by a cladding of $n=1$. The expression for $t$ is then the thickness of the slab for the lowest-order TE mode of a slab waveguide with effective refractive index $n_\text{eff}$\cite{Inan00}:
\begin{align}
t &= \frac{2}{\beta}\tan^{-1}\frac{\alpha}{\beta},\label{t} \quad\text{where} \\
\alpha &= \sqrt{k_x^2 - \left(\frac{\omega}{c}\right)^2}, \quad\text{and} \\
k_x &= \sqrt{\left(\frac{n_\text{eff}\omega}{c}\right)^2 - \beta^2}.
\end{align}
Here, $\alpha$ is the wave-vector which determines the evanescent decay of the slab waveguide mode into air, while $k_x$ is an approximated in-plane k-vector. To determine $n_\text{eff}$, the effective refractive index, we use the following approximation,
\begin{equation}
n_\text{eff} = \sqrt{\frac{\|\epsilon E^2\|}{\|E^2\|}}.
\end{equation}
% In the above formulation and for the remainder of the manuscript, the norm ($\|\cdot\|$) refers to the 2-norm over the entire grid and is equivalent to the spatial integral $(\int | \cdot |^2 dr)^{1/2}$.

>>>>

Further thought led us to conclude that
    the fundamental problem in the examples 
    in chapters \ref{direct} and \ref{regularized}
    is actually not in the methods themselves, 
    but in the improper selection of a target field. 
In fact, we realized thtat it is very difficult to select a suitable target field 
    because not every field even has a manufacturable dielectric structure 
    that is able to reproduce it. 
Furthermore, it is nearly impossible to select a multi-dimensional field 
    which corresponds to a well-behaved, isotropic and 
    discretely-valued $\epsilon$, 
    as would be needed for practical structures.

@@

\section{Results}
\subsection{1D}
Fig.~\ref{comp pic} shows that the iterative algorithm, after $400$ iterations and with the correct choice of regularization parameters $\eta_0$ and $\eta_1$, 
    indeed results in a well-behaved structure that is able to closely reproduce the modified target field\cite{Lu10}.

@@

\subsection{Iterative design in 2D}
The iterative design method can also be extended to two dimensions\cite{Lu10}.
We demonstrate this using an S-shaped target field 
    which is non-trivial to reproduce. 
The design result is shown in fig.~\ref{S pic}. 
    The resulting dielectric structure is continuous, unbounded and 
    contains very few singularities (white dots), 
    but the final target and actual fields match up well. 
Also, the computational cost remains quite reasonable; the $50$ iterations needed required only $5$ minutes of computation time. 
\begin{figure}\centering
\includegraphics[width=\textwidth]{p1/S}
\caption{Iterative design of an ``S'' resonator. 
The design was initialized by specifying an initial dielectric structure ($\epsilon=1$ everywhere) and a resonant field in the shape of an ``S''. 
The final dielectric structure was produced after $50$ iterations which took $90$ seconds to complete in total. The grid size was $80\times 120$. The final dielectric structure is quite unintuive, and yet reproduces the target field surprisingly well. This example demonstrates the versatility of the iterative design method in producing designs, from very simple specifications, which otherwise could be attained only with considerable difficulty.}
\label{S pic}
\end{figure}

@@

% \documentclass[10pt,letterpaper]{article}
% \usepackage{opex3}
% \usepackage{amsmath}
% \usepackage{graphicx}
% \begin{document}
% \title{Inverse design of a three-dimensional nanophotonic resonator}
% \author{Jesse Lu, Stephen Boyd, and Jelena Vu\v{c}kovi\'{c}}
% \address{Stanford University, \\ Stanford, C.A., 94305}
% \email{jesselu@stanford.edu}
% \begin{abstract} 
% The inverse design of a three-dimensional nanophotonic resonator is presented. The design methodology is computationally fast (10 minutes on a standard desktop workstation) and utilizes a 2.5-dimensional approximation of the full three-dimensional structure. As an example, we employ the proposed method to design a resonator which exhibits a mode volume of $0.32(\lambda/n)^3$ and a quality factor of $7063$. % 8808 35648
% \end{abstract}
% \ocis{(230.5750) Resonators.} % REPLACE WITH CORRECT OCIS CODES FOR YOUR ARTICLE
% %%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{thebibliography}{99}
% \bibitem{miller} D. A. B. Miller, ``Rationale and challenges for optical interconnects to electronic chips,'' Proc. of the IEEE \textbf{88}, 728-749 (2000).
% \bibitem{prevwork} J. Lu and J. Vuckovic, ``Inverse design of nanophotonic structures using complementary convex optimization,'' Opt. Express \textbf{18}, 3793-3804 (2010).
% \bibitem{inan} U. Inan, A. Inan, \emph{Electromagnetic Waves} (Prentice Hall, 2000), page 296.
% \bibitem{yee} K. Yee, ``Numerical solution of initial boundary value problems involving maxwell's equations in isotropic media,'' IEEE Trans. Antennas Propag. Mag. \textbf{14}, 302-307 (1966).
% \bibitem{boydbook} S. Boyd and L. Vandenberghe, \emph{Convex Optimization} (Cambridge University Press, 2004).
% \bibitem{altdir} S. Boyd, N. Parikh, E. Chu, B. Peleato, and J. Eckstein are preparing a manuscript to be called, ``Distributed Optimization and Statistical Learning via the Alternating Direction Method of Multipliers,'' \verb+http://www.stanford.edu/~boyd/papers/distr_opt_stat_learning_admm.html+. 
% \bibitem{cholmod} Y. Chen, T. A. Davis, W. W. Hager, and S. Rajamanickam, ``Algorithm 887: CHOLMOD, supernodal sparse Cholesky factorization and update/downdate,'' ACM Trans. Math. Software \textbf{35}, No. 3, 2009.
% \bibitem{cvx} M. Grant and S. Boyd, \emph{CVX: Matlab software for disciplined convex programming}, version 1.21. \verb+http://cvxr.com/cvx+, January 2011.
% \bibitem{digitize} D. Englund, I. Fushman, and J. Vuckovic, ``General recipe for designing photonic crystal cavities,'' Opt. Express \textbf{13}, 5961-5975 (2005).
% \end{thebibliography}
% %%%%%%%%%%%%%%%%%%%%%%%%%%  body  %%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Begin of 2.5 D (Motivation)}
% To date, the design of nanophotonic devices has generally involved a lengthy (days, weeks) process in which one perturbs, by trial-and-error, canonical structures such as photonic crystals or waveguide-coupled rings to achieve the desired performance for the device. 
% Instead, an inverse design method in which the user only specifies the desired electromagnetic field, or characteristics thereof, and then leaves the computer to find a dielectric structure satisfying these requirements, would be a much more intuitive and computationally efficient design strategy. Furthermore, such a method may even be the only feasible option available for the design of more complex multi-wavelength and multi-functional nanophotonic devices needed for on-chip integration\cite{miller}.
\subsection{Iterative design in 2.5D}
% \section{2.5-dimensional approximation}
Unfortunately, to directly extend our previous method to full three-dimensional space requires solving for a very large ($\sim 10^7 \times 10^7$) and ill-conditioned matrix, namely that given by the time-harmonic Maxwell's equation in three dimensions, which for the electric ($E$) field is
\begin{align}
\nabla\times\nabla\times E - \mu\omega^2\epsilon E &= 0,\quad\text{where}\label{wave} \\
E &= E(x,y,z).
\end{align}

@@

In this approximation, converting from a 2.5-dimensional structure to a full three-dimensional structure only requires the correct choice of slab thickness, $t$, since the in-plane values of $\epsilon$ remain the same. To estimate $t$ for the full three-dimensional device, we treat the resonant mode as a guided mode in a slab waveguide surrounded by a cladding of $n=1$. The expression for $t$ is then the thickness of the slab for the lowest-order TE mode of a slab waveguide with effective refractive index $n_\text{eff}$\cite{Inan00}:
\begin{align}
t &= \frac{2}{\beta}\tan^{-1}\frac{\alpha}{\beta},\label{t} \quad\text{where} \\
\alpha &= \sqrt{k_x^2 - \left(\frac{\omega}{c}\right)^2}, \quad\text{and} \\
k_x &= \sqrt{\left(\frac{n_\text{eff}\omega}{c}\right)^2 - \beta^2},
\end{align}
Here, $\alpha$ is the wave-vector which determines the evanescent decay of the slab waveguide mode into air, while $k_x$ is an approximated in-plane k-vector, and $x$ is the speed of light. To determine $n_\text{eff}$, the effective refractive index, we use the following approximation,
\begin{equation}
n_\text{eff} = \sqrt{\frac{\|\epsilon E^2\|}{\|E^2\|}}.
\end{equation}
% In the above formulation and for the remainder of the manuscript, the norm ($\|\cdot\|$) refers to the 2-norm over the entire grid and is equivalent to the spatial integral $(\int | \cdot |^2 dr)^{1/2}$.

### 2013-05-24 thesis/chapter3.tex
Thesis revision (from Jelena's comments) almost done.
Just need conclusion.

\chapter{Objective-first design} \label{ob-1}

@@

Specifically, we present five results which suggest our method
    is indeed able to design arbitrary single-mode, linear nanophotonic devices\cite{Lu12}.
These results comprise of couplers between 
    the following nanophotonic waveguides:
    \begin{enumerate}
    \item coupling between waveguides of different refractive index and width
        (Fig.~\ref{fig:fiber});
    \item coupling between waveguide modes of different order and symmetry
        (Fig.~\ref{fig:mode});
    \item coupling between waveguides that confine light 
        using different principles 
        (index guided vs. distributed Bragg reflection guided), 
        i.e., between a slab waveguide and a photonic crystal fiber
        (Fig.~\ref{fig:aircore})
    \item coupling from a dielectric to a plasmonic metal-insulator-metal 
        waveguide (Fig.~\ref{fig:mim}); and
    \item coupling from a dielectric waveguide to a (plasmonic) metal wire
        (Fig.~\ref{fig:wire}).
    \end{enumerate}
All the designs presented are both highly efficient 
    ($\sim 98\%$ coupling efficiency, in contrast to previous designs
    which achieve only $\sim 70\%$ coupling efficiency\cite{Veronis07,Yang10})
    and extremely compact 
    (with footprints of only 1.5-4 square vacuum wavelengths).

>>>>

\chapter{Objective-first design in 2D} \label{ob-1}

@@

Specifically, we present five results which suggest our method
    is indeed able to design arbitrary single-mode, linear nanophotonic devices\cite{Lu12}.
These results comprise of couplers between 
    the following nanophotonic waveguides:
    \begin{enumerate}
    \item coupling between waveguides of different refractive index and width
        (Fig.~\ref{fig:fiber});
    \item coupling between waveguide modes of different order and symmetry
        (Fig.~\ref{fig:mode});
    \item coupling between waveguides that confine light 
        using different principles 
        (index guided vs. distributed Bragg reflection guided), 
        i.e., between a slab waveguide and a photonic crystal fiber
        (Fig.~\ref{fig:aircore})
    \item coupling from a dielectric to a plasmonic metal-insulator-metal 
        waveguide (Fig.~\ref{fig:mim}); and
    \item coupling from a dielectric waveguide to a (plasmonic) metal wire
        (Fig.~\ref{fig:wire}).
    \end{enumerate}
All designs presented are both highly efficient 
    ($\sim 98\%$ coupling efficiency, in contrast to previous designs
    which achieve only $\sim 70\%$ coupling efficiency\cite{Veronis07,Yang10})
    and extremely compact 
    (with footprints of only 1.5-4 square vacuum wavelengths).

### 2013-05-24 thesis/chapter4.tex
Thesis revision (from Jelena's comments) almost done.
Just need conclusion.

\chapter{Nanophotonic computational design} \label{final}

@@

We now develop a method which may indeed be capable of designing 

@@

All of our results are in three dimensions
    and are planar structures, consisting of a 250 nm etched silicon slab
    completely surrounded by silica.
The permittivity values of silicon and silica used
    were $\epsilon_\text{Si} = 12.25$ and $\epsilon_\text{SiO$_2$} = 2.25$
    respectively.

@@

\myfig{3Db_wgc_te2}
    {Perfomance specification of the TE mode converter.
    Input mode is the fundamental TE-polarized mode on the left.
    Primary output mode is the second-order mode on the right.
    Output power in the transmitted fundamental mode 
    should be no more than 1\%.
    The structure shown is the final three-dimensional design
    (the same holds for all the following figures in the article).
%     The x- and y-axes are in the plane of the structure,
%         with the x-axis is the direction of propagation 
%         for the input and output waveguides.
%     As expected, the z-axis is perpendicular to the plane of the structure.
    }

@@

\myfig{3Db_wgc_tm}
    {Perfomance specification of the TM mode converter.
    Input mode is the fundamental TM-polarized mode on the left.
    Primary output mode is the second-order mode on the right.
    Output power in the transmitted fundamental mode on the right 
    above 1\% is to be rejected.
    The structure shown is the final three-dimensional design.}

@@

\myfig{3Db_hub_4x4}
    {4$\times$4 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode.
    Output power into the desired output arm is specified to be greater than 90\%,
        no rejection modes are used for computational efficiency.
    This hub directs input power from input ports 1, 2, 3, and 4
    into output ports 3, 2, 4, and 1 respectively.}
\myfig{hub_4x4_comb}
    {4$\times$4 hub final result.
    The conversion efficiencies into the selected output arms
        are 85.9\%, 88.1\%, 85.4\%, and 84.3\% for
        input arms 1, 2, 3, and 4 respectively (top to bottom).}

@@

Our performance specification (\fig{3Db_hub_2x2x2}) is
    to cross-couple the waveguides at the 1310 nm wavelength,
    but to uncouple the waveguides at the 1550 nm wavelength.
The final result is shown in \fig{hub_2x2x2}.
\myfig{3Db_hub_2x2x2}
    {2$\times$2$\times$2 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode
        at either 1550 nm or 1310 nm wavelengths.
    Output power into the desired output arm is specified to be greater than 80\%,
        no rejection modes are used for computational efficiency.
    This hub directs input arms 1 and 2 into output arms 1 and 2 at 1550 nm,
        but swaps them at 1310 nm.}
\myfig{hub_2x2x2}
    {2$\times$2$\times$2 hub final result.
    The conversion efficiencies at the 1550 nm wavelength 
        are 77.6\% and 73.7\% respectively for the top and bottom inputs.
    At 1310 nm, the respective efficiencies are 75.7\% and 75.2\%.}

>>>>

\chapter{Nanophotonic computational design in 3D} \label{final}

@@

In this central chapter of this work, 
    we present a method which may indeed be capable of designing 

@@

\emph{All of our results are in three dimensions
    and are planar structures, consisting of a 250 nm etched silicon slab
    completely surrounded by silica.}
The permittivity values of silicon and silica used
    were $\epsilon_\text{Si} = 12.25$ and $\epsilon_\text{SiO$_2$} = 2.25$
    respectively.

@@

\myfig{3Db_wgc_te2}
    {Perfomance specification of the TE mode converter.
    Input mode is the fundamental TE-polarized mode on the top-left.
    Primary output mode is the second-order mode on the top-right.
    Output power in the transmitted fundamental mode (bottom-right) 
    should be no more than 1\%.
    The structure shown is the final three-dimensional design
    (the same holds for all the following figures in the article).
%     The x- and y-axes are in the plane of the structure,
%         with the x-axis is the direction of propagation 
%         for the input and output waveguides.
%     As expected, the z-axis is perpendicular to the plane of the structure.
    }

@@

\myfig{3Db_wgc_tm}
    {Perfomance specification of the TM mode converter.
    Input mode is the fundamental TM-polarized mode on the top-left.
    Primary output mode is the second-order mode on the top-right.
    Output power in the transmitted fundamental mode on the bottom-right 
    should be under 1\%.
    The structure shown is the final three-dimensional design.}

@@

\myfig{3Db_hub_4x4}
    {4$\times$4 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode.
    Output power into the desired output arm is specified to be greater than 90\%,
        no rejection modes are used for computational efficiency.
    This hub directs input power from input ports 1, 2, 3, and 4
    into output ports 3, 2, 4, and 1 respectively.}
\myfig{hub_4x4_comb}
    {4$\times$4 hub final result.
    The conversion efficiencies into the selected output arms
        are 85.9\%, 88.1\%, 85.4\%, and 84.3\% for
        input arms 1, 2, 3, and 4 respectively (top to bottom).}
\clearpage

@@

Our performance specification (\fig{3Db_hub_2x2x2}) is
    to cross-couple the waveguides at the 1310 nm wavelength,
    but to uncouple the waveguides at the 1550 nm wavelength.
The final result is shown in \fig{hub_2x2x2}.
\myfig{3Db_hub_2x2x2}
    {2$\times$2$\times$2 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode
        at either 1550 nm or 1310 nm wavelengths.
    Output power into the desired output arm is specified to be greater than 80\%,
        no rejection modes are used for computational efficiency.
    This hub directs input arms 1 and 2 into output arms 1 and 2 at 1550 nm,
        but swaps them at 1310 nm.}
\myfig{hub_2x2x2}
    {2$\times$2$\times$2 hub final result.
    The conversion efficiencies at the 1550 nm wavelength 
        are 77.6\% and 73.7\% respectively for the top and bottom inputs.
    At 1310 nm, the respective efficiencies are 75.7\% and 75.2\%.}
\clearpage

### 2013-05-24 thesis/intro.tex
Thesis revision (from Jelena's comments) almost done.
Just need conclusion.

Currently, on-chip networks consist of devices
    which are designed by tuning
    a handful of design parameters (\fig{integrated_circuit}).
Much progress has been made on this front;
    however, we note that the available design space is largely untouched (\fig{des_complexity}).

\myfig{integrated_circuit}
    {Examples of hand-tuned nanophotonic devices.
    Device performance is tuned by varying the sizes of and spacing between
    a handful of elements (waveguides, resonators, gratings, \ldots).}

To be precise, the motivation for this works stems from the twin realizations that
    \begin{enumerate}
    \item the fabrication of on-chip optical networks in semiconductor 
    foundries enables virtually unlimited design complexity at no additional 
    fabrication cost, and
    \item component performance can only benefit by the addition and
    consideration of these additional degrees of freedom.
    \end{enumerate}
The goal of this work, then, is to be able to take advantage
    of the full available parameter space to design nanophotonic devices.

@@

Along this line of reasoning
    we expect to be able to produce devices which
    \BI have very compact footprints,
    \I  exhibit high efficiency performance,
    \I  are multi-modal in their operation,
    \I  are robust to shifts in wavelength and temperature,
    \I  are robust to fabrication error,
    \I  may even demonstrate novel, never-before-seen functionality. \EI
Our final result (chapter~\ref{final}) indeed demonstrates
    manufacturable, three-dimensional nanophotonic devices which
    exhibit these properties.

Of course, in this regime, the hand-tuning of parameters 
    is no longer a tractable strategy.
Instead, we use a \emph{design-by-specification} strategy
    which simply means that our software produces nanophotonic designs
    based solely based on the desired performance specification given to it,
    and requires no intermediate hand tuning.

We note, lastly, that the design-by-specification strategy
    is unusually powerful in the sense that \emph{all} useful nanophotonic devices
    can be designed in this way.
Indeed, the apparent ease with which we are able to produce designs seems
    to lend one to this conclusion.

@@

>>>>

Currently, on-chip networks consist of devices
    which are designed by tuning
    a handful of design parameters\cite{pic1,pic2} (\fig{integrated_circuit}).
Much progress has been made on this front;
    however, we note that the available design space is largely untouched (\fig{des_complexity}).

\myfig{integrated_circuit}
    {Examples of hand-tuned nanophotonic devices\cite{pic1,pic2}.
    Device performance is tuned by varying the sizes of and spacing between
    a handful of elements (waveguides, resonators, gratings, \ldots).}

To be precise, the motivation for this works stems from the twin realizations that
    \begin{enumerate}
    \item the fabrication of on-chip optical networks in semiconductor 
    foundries enables virtually unlimited design complexity at no additional 
    fabrication cost, and
    \item component performance can only benefit by the addition and
    consideration of these additional degrees of freedom.
    \end{enumerate}
Unfortunately, the lack of intuition for what such designs might look like and
    the inability to manually search such a large parameter space
    have greatly hindered the ability to achieve a full-parameter space design.
The goal of this work, then, is to be able to take advantage
    of the full available parameter space to design nanophotonic devices.

@@

Previously, design algorithms featuring a large number of parameters
    have been applied to the design of individual devices.
These range from the use of genetic algorithms\cite{lipson},
    level-set methods\cite{yablo}, 
    and the optimization of specific geometric parameters\cite{levi,johnson}
    to design nanophotonic devices.
Additionally, much success in this field has been attained\cite{sigmund1}
    through the use of topological design methods that
    have traditionally been applied in other fields \cite{jameson,bendsoe}.

Built upon this prior art,
    this work's aim is to produce a design method 
    that not only produces designs for a single device,
    but for a wide range of devices.
Specifically,
    we desired to be able to produce devices which
    \BI have very compact footprints,
    \I  exhibit high efficiency performance,
    \I  are multi-modal in their operation,
    \I  are robust to shifts in wavelength and temperature,
    \I  are robust to fabrication error,
    \I  may even demonstrate novel, never-before-seen functionality. \EI
Our final result (chapter~\ref{final}) indeed demonstrates
    manufacturable, three-dimensional nanophotonic devices which
    exhibit these properties.

Not only was a large parameter space used to achieve such designs,
    but we employed a \emph{design-by-specification} strategy
    which simply means that our software produces nanophotonic designs
    based solely on the desired performance specification given to it,
    and requires no intermediate hand tuning.
We note that such a design-by-specification strategy
    is unusually powerful in the sense that \emph{all} useful nanophotonic devices
    can be designed in this way.
Indeed, the apparent ease with which we are able to produce designs seems
    to lend one to this conclusion.

@@

The work presented here is based on the following publications:
\begin{itemize}
\item chapters~\ref{direct}-\ref{bounded}: ``Inverse design of nanophotonic structures using complementary convex optimization,'' Opt. Express (2010),
\item chapter~\ref{bounded}:
    ``Inverse design of a three-dimensional nanophotonic resonator,''
    Opt. Express (2011), 
\item chapter~\ref{ob-1}: ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes,'' 
    Opt. Express (2012),
\item chapter~\ref{final}: ``Nanophotonic computational design,'' Opt. Express (2013).
\end{itemize}
